---
title: "Internal Function"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following functions are used internally to the main functions:

## Find in harbour
<br>

```{r funcdef8, echo = F, eval = F}
find_inport(data, ports)
```

```{r funcdef8, eval = F}
```
<br>

* data:
* ports:

## Find the closest harbour
This function was used to assign the harbor to fishing session the start or finish without a "in harbor" position. The departure or the arrival harbor was assigned considering the closest harbor respect to the first or last position of the session:
<br>

```{r funcdef1, echo = F, eval = F}
closest_port(longitude , latitude, ports)
```

```{r funcdef1, eval = F}
```
<br>

* longitude-latitude: are the coordinates of the first or last point of the fishing session
* ports: is the shapefile with harbors locations

## Import parameters
<br>

```{r funcdef3, echo = F, eval = F}
import_parameters(parameters, centroids)
```

```{r funcdef3, eval = F}
```
<br>

* parameters: an external csv file with the following parameters:
  

* centroids: an external csv file containing the set of centroids values to test using the kmeans method in the classification alghorithm

## Classfied speed
<br>

```{r funcdef4, echo = F, eval = F}
classify_speed(data, gear, xcentroids, pars)
```

```{r funcdef4, eval = F}
```
<br>

* data:
* gear
* xcentroids
* pars

## Make segment
<br>

```{r funcdef5, echo = F, eval = F}
make_segment_lite(data, gear)
```

```{r funcdef5, eval = F}
```
<br>

* data:
* gear:

## Assign ping to session
<br>

```{r funcdef6, echo = F, eval = F}
assign_session(data, session_table)
```

```{r funcdef6, eval = F}
```
<br>

* data:
* session_table:


## Search cluster
<br>

```{r funcdef7, echo = F, eval = F}
search_cluster(data, pars, gear)
```

```{r funcdef7, eval = F}
```
<br>

* data:
* pars:
* gear:

## Check cluster
<br>

```{r funcdef9, echo = F, eval = F}
check_cluster(data, gear, threshold, low_speed)
```

```{r funcdef9, eval = F}
```
<br>

* data: 
* gear:
* threshold:
* low_speed:


## Data partition
This is the function used to split a validated dataset in validation, training and test subset. The input dataset is represented by a validate 
<br>

```{r funcdef2, echo = F, eval = F}
data_partitioning(data)
```

```{r funcdef2, eval = F}
```
<br>

* data: it is an input dataframe containing the ID of the vessel, the session and the validated gear


## Data partition
This is the function used to split a validated dataset in validation, training and test subset. The input dataset is represented by a validate 
<br>

```{r funcdef999, echo = F, eval = F}
identify_trasmission_gaps(data, coord_sys)
```

```{r funcdef999, eval = F}
```
<br>

* data: it is an input dataframe containing the ID of the vessel, the session and the validated gear